---
title: "Missing Values"
format: 
  html:
    embed-resources: true
editor: visual
---

## Missing Values

```{r}
library(knitr)
library(tidyverse)
```

```{r}
#First let's cast the Value column as numeric values:

straw_cen$Value<-as.numeric(str_replace(straw_cen$Value,",",""))
```

## Types of Missingness:

### Missing Completely at Random:

A variable is missing completely at random if the probability of missingness is the same for all units.

### Missing at Random:

Probability of missingness depends on data. For strawberry census, an example is to look at the values that are NA and see they depend on state or county.

### Missingness that depends on unobserved predictors
Information not in data

### Missingness that depends on the missing data:
E.g we have earnings for some group of people. Over time that change

How do we visualize this? First, note that where County is NA but state is not, we have state-wide figures. E.g:

```{r}
agg_calif<-straw_cen |> filter(State=="CALIFORNIA") |> filter(is.na(County))
```

Now, we'll create a dataframe that has all the missing values

```{r}
missing<-straw_cen |> filter(is.na(Value)) |> select(c(`Geo Level`,State,County,Measure:Value))
missing |> distinct(`Geo Level`)
print(missing |> distinct(State),n=60)
missing |> distinct(County)
```

Clearly there is a LOT of missingness at the county level. There are 1006 distinct counties in the dataset and 786 of them have missing data.

What about Missingness at a State Level?

```{r}
state_level<-straw_cen |> filter(`Geo Level` == "STATE")
dim(state_level |> filter(is.na(Value)))[1]/dim(state_level)[1]
```
Not bad. Take a look at the data
```{r}
state_level_missingness<-state_level |> filter(is.na(Value))
state_level_missingness |> distinct(size_bracket)
```

### Strategies for imputation:

General rule of thumb: If you can fill in the blanks with data that is more granular than your missing value, i.e if you can 'reconstruct' or re-caclulate something, that's always better than imputing downwards.

For example:
```{r}
california<-straw_cen |> filter(State=="CALIFORNIA")
state_level_calif<-state_level |> filter(State=="CALIFORNIA")
california_acres_bearing<- california |> filter(Measure == "ACRES" & Bearing_type=="BEARING")


dim(california |> filter(Measure == "ACRES" & Bearing_type=="BEARING") |> filter(is.na(Value)))[1]/dim(california |> filter(Measure == "ACRES" & Bearing_type=="BEARING"))[1]
```


If that is unavoidable however, which it is likely to be in most of cases, a few methods include:

* Averages
* Proportional Adjustment
* Regression

This isn't an exhaustive list by any stretch of the imagination but a good starting point. 

Example with california data:
Consider acres_bearing:

```{r}
#We can start by just taking the difference between the total and the data we have and splitting it equally between the two:

acres_bearing_total<-state_level_calif |> filter(Measure=="ACRES" &
                                                 Bearing_type=="BEARING" &
                                                 Domain == "TOTAL") |> 
                                                 select(Value)
acres_bearing<-state_level_calif |> filter(Measure=="ACRES" &
                                           Bearing_type=="BEARING" & !is.na(Value)
                                           & Domain != "TOTAL")

imputed_val<-(acres_bearing_total -sum(acres_bearing$Value))/2

```
This isn't efficient code at all. Can you write a function that does this?

Note also however, that we can use the Area Grown Column. What information does that give us?

How do we impute from this? Do we just split it in 2 and assign it to bearing and non-bearing. Do we calculate the total across both and split in proportion:
```{r}
calif_acres<-state_level |> filter(State=="CALIFORNIA"& Measure=="ACRES")
bearing_vec<-calif_acres |> filter(Bearing_type=="BEARING") 
bearing_vec<-bearing_vec$Value
bearing_indices<-!is.na(bearing_vec)
grown_vec<-calif_acres |> filter(Bearing_type=="GROWN") 
grown_vec<-grown_vec[bearing_indices,]$Value

bearing_vec_noNA<-bearing_vec[bearing_indices]

non_bearing_vec<-calif_acres |> filter(Bearing_type=="NON-BEARING")

mean(bearing_vec_noNA/grown_vec)

```

Alternatively, run a Linear Regression:
```{r}

#Filter out the NA Values
state_level_noNA<-state_level |> filter(!is.na(Value))

fit1<-lm(Value~State+Measure+Bearing_type+Domain+size_bracket,data=state_level_noNA)

```

