---
title: "Chemical Usage in Strawberry Farming - EDA Report"
author: "Your Name"
date: "2024-10-21"
format:
  html:
    code-fold: true
editor: visual
---

# Introduction

This exploratory data analysis (EDA) aims to understand the usage of chemicals in strawberry farming in the United States. We will utilize both internal data (census, survey, and organic strawberry datasets) and external sources to provide context, examine toxicity, and evaluate chemical impacts on farming yield and safety. The focus is on extracting insights that could answer significant questions and raise further areas of investigation.

The data provided include:

-   **Survey Data**: Survey-based strawberry production metrics.
-   **Census Data**: Comprehensive census data on strawberry production.
-   **Organic Data**: Census-based information on organic strawberries.
-   **Chemical Data**: Information regarding the chemicals used in strawberry farming.

In this EDA, we will begin by exploring the chemical data, identifying key trends, and guiding our analysis towards specific questions regarding chemical toxicity, hazards, and impacts on production.

# Load Libraries and Data

```{r}
# Loading necessary libraries
library(tidyverse)
library(knitr)
library(kableExtra)

# Load the datasets
#census_data <- read_csv("./_census_data.csv")
#organic_data <- read_csv("/mnt/data/_census_organic_data.csv")
#chemical_data <- final_chemical_data
#survey_data <- read_csv("/mnt/data/_survey_data.csv")
```

# Data Overview

Let's begin by gaining an overview of the chemical data. We will start by examining the columns available, identifying unique chemicals, and summarizing their usage in strawberry farming.

## Chemical Data Overview

```{r include=false}
# Loading necessary libraries
library(tidyverse)
library(knitr)
library(kableExtra)

```

## Yearly Trends in Chemical Application

```{r}
# Summarize the chemical application by year
chemical_trend <- chemical_data %>%
  group_by(Year) %>%
  summarize(total_chemicals = n()) %>%
  arrange(Year)

# Plot the yearly trends in chemical application
ggplot(chemical_trend, aes(x = Year, y = total_chemicals)) +
  geom_line(color = "steelblue", size = 1) + 
  geom_point(color = "darkblue", size = 3) + 
  labs(title = "Yearly Trends in Chemical Application",
       x = "Year",
       y = "Total Chemical Applications") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) # Centering the title

```

The line plot above visualizes the **yearly trends in chemical applications** in strawberry farming from **2018 to 2023**. The data reveals a few key points:

1.  **2019** saw the **highest number of chemical applications**, with over 900 instances.

2.  Following this peak, there was a **significant drop** in chemical applications by **2021**, bringing the total down to just above 850.

3.  Despite the decline, there was a **slight increase in 2023**, which may indicate a shift back towards higher usage of chemicals.

The observed dip in chemical applications from 2019 to 2021 could likely be attributed to the global disruptions caused by the COVID-19 pandemic. Supply chain issues, labor shortages, and financial constraints may have limited farmers' access to chemical inputs during this time. Additionally, the pandemic brought increased awareness of sustainability and health, potentially driving a shift toward reduced chemical usage in favor of organic or eco-friendly farming practices. The subsequent recovery in chemical applications by 2023 might reflect the normalization of supply chains and agricultural operations as the world adjusted to post-pandemic conditions.

```{r}
# Summarize the types of chemicals and their occurrences
chemical_summary <- chemical_data %>%
  group_by(type, use) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

# View the summary
#chemical_summary

# Create a bar chart with count labels
ggplot(chemical_summary, aes(x = use, y = count, fill = use)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_text(aes(label = count), position = position_dodge(width = 0.9), vjust = -0.5) +
  theme_minimal() +
  labs(title = "Chemical Use Distribution by Type (2018-2023)",
       x = "Chemical Use",
       y = "Count") +
  scale_fill_brewer(palette = "Set2")
```

## Top 5 Most Common Chemicals Used

```{r}
# Filter the data for the years 2018 to 2023
chemical_data_filtered <- chemical_data %>%
  filter(Year >= 2018 & Year <= 2023, name != "TOTAL)")

# Calculate the total usage of each chemical from 2018 to 2023
top_chemicals_usage <- chemical_data_filtered %>%
  group_by(name) %>%
  summarise(
    total_usage = sum(Value, na.rm = TRUE),
    use = first(use)  # Assuming each chemical has a unique `type`
  ) %>%
  arrange(desc(total_usage)) %>%
  slice(1:5) # Select the top 5 chemicals

# Create a plot for the top 5 chemicals
ggplot(top_chemicals_usage, aes(x = reorder(name, -total_usage), y = total_usage)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  geom_text(aes(label = round(total_usage, 1)), vjust = -0.3, size = 3.5) + # Adding text labels
  labs(title = "Top 5 Chemicals Used from 2018 to 2023",
       x = "Chemical Names",
       y = "Total Usage (kg)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5))

# Check for increasing trends in chemical usage over time
usage_trends <- chemical_data_filtered %>%
  filter(name %in% top_chemicals_usage$name) %>%
  group_by(Year, name) %>%
  summarise(yearly_usage = sum(Value, na.rm = TRUE)) %>%
  ungroup()

# Create a line plot to show the trends over time for the top 5 chemicals
ggplot(usage_trends, aes(x = Year, y = yearly_usage, color = name)) +
  geom_line(size = 1) +
  theme_minimal() +
  labs(
    title = "Usage Trends of Top 5 Chemicals from 2018 to 2023",
    x = "Year",
    y = "Yearly Usage (kg)",
    color = "Chemical Name"
  )

```

The analysis of chemical usage trends in strawberry farming has revealed that **Chloropicrin** is the most heavily utilized chemical, with a marked increase in its use from 2018 to 2023. The next step in our exploratory data analysis (EDA) is to dive deeper into understanding what these chemicals are, focusing on:

-   **What do these chemicals do exactly?** Understanding their purpose in strawberry farming, such as whether they are fungicides, pesticides, or fumigants.

-   **Are they harmful?** Evaluating the potential health and environmental risks associated with these chemicals.

-   **Why has there been a significant increase in the use of Chloropicrin since 2018?** Investigating the factors driving this increase, whether it is related to changes in regulations, pest pressures, or shifts in agricultural practices.

-   **What exactly is Chloropicrin?** Gaining a detailed understanding of Chloropicrin, including its properties, application methods, and effectiveness.

## Chloropicrin.

Let's see what chemical type Chloropicrin belongs to.

```{r}
print(top_chemicals_usage %>%
  filter(name == "CHLOROPICRIN") %>%
  select(use))
```

Well, it appears that our current dataset will not be able help us identifying Chloropicrin's chemical use. Fortunately, we have its chemical code value and will use external datasets, such as the `EPA Toxic Chemical Dashboard` and PubChem to provide detailed information on chemical properties, toxicity, and usage guidelines.

These external resources will help us interpret the implications of using these chemicals and assess the potential risks and benefits of their application in strawberry farming. This analysis will guide us towards better decision-making in the context of agricultural practices and sustainability.

```{r}
# Extracting the chemical code of Chloropicrin
print(chemical_data %>%
  filter(name == "CHLOROPICRIN") %>%
  select(code) %>%
  head(1)) 
```

Which state uses Chloropicrin more?

```{r}
print(chemical_data %>%
        filter(name == "CHLOROPICRIN") %>%
          total_usage = sum(Value, na.rm = TRUE),
      )
```
