## Handling Missing Values with State-Level Ratios

The following functions have been replaced with a more optimized imputation strategy that incorporates both interpolation and scaling of imputed values to ensure accuracy.

### Interpolation and Scaling for Missing Values

The goal of this section is to efficiently handle missing values in the `census_data` dataframe. The functions in this section perform the following:

1. **Interpolation of Missing Values**: Using logical assumptions and calculations to impute missing data points where either the `BEARING`, `NON-BEARING`, or `GROWN` values are not available.
2. **Scaling of Imputed Values**: Ensuring that after imputation, the newly generated values are consistent with the totals reported for each state and measure.
3. **Handling Different Scenarios**: Differentiating between cases where both `BEARING` and `NON-BEARING` values are missing versus only one is missing.
4. **Maintaining Data Consistency**: Applying proportional adjustments to ensure that the calculated values add up to reported totals, thereby maintaining consistency and accuracy of the dataset.

```{r}
#| label: explore census_data$Domain & Domain Category
#| echo: false

## remove AREA GROWN and parens 
## change NOT SPECIFIC TO TOTAL

census_data <- census_data |> rename(size_bracket = `Domain Category`)

census_data$size_bracket <- str_replace(census_data$size_bracket, "NOT SPECIFIED", "TOTAL")

census_data$size_bracket <- str_replace(census_data$size_bracket, "AREA GROWN: ", "")


#We can deal with those cases as follows:
#We are going to write a function that does the imputation for one index
#Then we can just sapply it over the vector if na indices and finally
#assign that to the missing values.

#First we adapt our previous proportion function to size brackets as a prop of total
#This is a function we shall use below


prop_by_size_bracket <- function(state, measure, bearing, bracket) {
  state_cen <- census_data |> filter(State == state & Measure == measure)
  bracket_vec <- state_cen |> filter(size_bracket == bracket)
  bracket_vec <- bracket_vec$Value
  bracket_indices <- which(!is.na(bracket_vec))
  
  if (length(bracket_indices) != 0) {
    total_vec <- state_cen |> filter(size_bracket == "TOTAL")
    total_vec <- total_vec[bracket_indices, ]$Value
    
    bracket_no_na <- which(!is.na(bracket_vec))
    total_no_na <- which(!is.na(total_vec))
    
    if (length(intersect(bracket_no_na, total_no_na)) != 0) {
      not_na_indices <- intersect(bracket_no_na, total_no_na)
      
      bracket_vec_noNA <- bracket_vec[not_na_indices]
      total_vec_noNA <- total_vec[not_na_indices]
      
    } else {
      bracket_vec_noNA <- mean(bracket_vec[bracket_no_na])
      total_vec_noNA <- mean(total_vec[total_no_na])
    }
    return(mean(bracket_vec_noNA / total_vec_noNA))
  } else{
    sum_non_total <- state_cen |> filter(size_bracket != "TOTAL") |> filter(!is.na(Value))
    sum_non_total <- sum(sum_non_total$Value)
    
    total <- state_cen |> filter(size_bracket == "TOTAL")
    total <- sum(total$Value)
    #There's a small technicality to deal with here:
    #The rest of this function is returning a proportion but this part is not.
    #It gives us the actual imputed value.
    #That's fine but we need to be careful how we use it in the final function
    #To do that, we'll return a vector with an arbitrary second element
    #Then we can check the type when we call this function and use it accordingly
    return(as.character((total - sum_non_total) / length(bracket_vec)))
  }
}

```

```{r}
#| label: imputation final
#| echo: false

state_vec <- census_data |> distinct(State)

# Retrieving indices of rows with NA Values
bearing_na_indices <- which(is.na(census_data$Value) &
                              census_data$bearing_type == "BEARING")
non_bearing_na_indices <- which(is.na(census_data$Value) &
                                  census_data$bearing_type == "NON-BEARING")
grown_na_indices <- which(is.na(census_data$Value) &
                            census_data$bearing_type == "GROWN")

# Function for calculating the proportion between bearing and grown
bearing_non_bearing_prop_avg <- function(state, measure, bearing) {
  state_cen <- census_data |> filter(State == state & Measure == measure)
  bearing_vec <- state_cen |> filter(bearing_type == bearing)
  bearing_vec <- bearing_vec$Value
  bearing_indices <- which(!is.na(bearing_vec))
  
  grown_vec <- state_cen |> filter(bearing_type == "GROWN")
  grown_vec <- grown_vec[bearing_indices, ]$Value
  
  bearing_no_na <- which(!is.na(bearing_vec))
  grown_no_na <- which(!is.na(grown_vec))
  
  not_na_indices <- intersect(bearing_no_na, grown_no_na)
  
  bearing_vec_noNA <- bearing_vec[not_na_indices]
  grown_vec_noNA <- grown_vec[not_na_indices]
  
  return(mean(bearing_vec_noNA / grown_vec_noNA))
}

# Imputation function for GROWN values
census_grown_imputer <- function(i) {
  state <- as.character(census_data[i, "State"])
  measure <- as.character(census_data[i, "Measure"])
  size_brack <- as.character(census_data[i, "size_bracket"])
  
  non_bearing_val <- as.numeric(
    census_data |> filter(
      State == state &
        Measure ==   measure &
        size_bracket == size_brack &
        bearing_type == "NON-BEARING"
    ) |> select(Value)
  )
  
  bearing_val <- as.numeric(
    census_data |> filter(
      State == state &
        Measure ==   measure &
        size_bracket == size_brack &
        bearing_type == "BEARING"
    ) |> select(Value)
  )
  
  # Handling different cases of missing values
  if (!is.na(non_bearing_val) & !is.na(bearing_val)) {
    return(bearing_val + non_bearing_val)
  } else if ((!is.na(non_bearing_val) & is.na(bearing_val))) {
    prop <- bearing_non_bearing_prop_avg(state, measure, bearing = "NON-BEARING")
    return(non_bearing_val / prop)
  } else if ((!is.na(bearing_val) & is.na(non_bearing_val))) {
    prop <- bearing_non_bearing_prop_avg(state, measure, bearing = "BEARING")
    return(bearing_val / prop)
  } else {
    bearing <- "GROWN"
    prop <- prop_by_size_bracket(state, measure, bearing, size_brack)
    total <- as.numeric(
      census_data |> filter(
        State == state &
          Measure == measure &
          size_bracket == "TOTAL" & bearing_type == bearing
      ) |> select(Value)
    )
    if (class(prop) == "character") {
      return(as.numeric(prop))
    } else {
      return(prop * total)
    }
  }
}

# Applying the imputation function to GROWN values
grown_imputed <- sapply(grown_na_indices, census_grown_imputer)
census_data$Value[grown_na_indices] <- grown_imputed

# Function to scale imputed values to ensure they sum to the total
imputed_val_scaler <- function(index_vec, bearing) {
  scaled_imputed_vals <- rep(0, length(index_vec))
  index_tracker <- c()
  for (i in index_vec) {
    if (!(i %in% index_tracker)) {
      grouped_indices <- index_vec[which(census_data$State[index_vec] == census_data$State[i])]
      state <- census_data$State[i]
      measure <- census_data$Measure[i]
      state_total_cen <- census_data |> filter(State == state &
                                               Measure == measure & bearing_type == bearing)
      state_total_cen_not_total <- state_total_cen |> filter(Domain != "TOTAL")
      overall_total <- state_total_cen |> filter(Domain == "TOTAL")
      overall_total <- overall_total$Value
      grouped_indices_adj <- grouped_indices - which(
        census_data$State == state &
          census_data$Measure == measure & census_data$bearing_type == bearing
      )[1] + 1
      correct_total <- overall_total - sum(state_total_cen_not_total$Value[-grouped_indices_adj])
      incorrect_total <- sum(state_total_cen_not_total$Value[grouped_indices_adj])
      scaled_vals <- (correct_total / incorrect_total) * state_total_cen_not_total$Value[grouped_indices_adj]
      scaled_imputed_vals[which(index_vec %in% grouped_indices)] <- scaled_vals
      index_tracker <- c(grouped_indices, index_tracker)
    }
  }
  return(scaled_imputed_vals)
}

grown_imputed_scaled <- imputed_val_scaler(grown_na_indices, bearing = "GROWN")
census_data$Value[grown_na_indices] <- grown_imputed_scaled
```